---
title: "Lab 1-2 Replication"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
always_allow_html: yes
---

## Introduction
This notebook provides a replication of Lab 1-2.

## Dependencies
This notebook requires the following packages:

```{r load-packages}
# tidyverse packages
library(ggplot2)       # static mapping

# mapping packages
library(mapview)      # preview spatial data
library(sf)           # spatial tools

# other packages
library(here)         # file path management
library(viridis)      # viridis color palettes
```

## Load Data
The data for today's lecture are county-level population estimates from the 2010 Decennial Census for Missouri. These can be found in `data/MO_DEMOS_Black_Population`. They are `.geojson` data:

```{r load-data}
mo <- st_read(here("data", "MO_DEMOS_Black_Population", "MO_DEMOS_Black_Population.geojson"))
```
We're now ready to map these data.

## Part 1 - Exploration
We'll use the `mapview()` function to explore these data:

```{r}
mapview(mo)
```

We can see that our data line-up with the basemap (always a good sign!) and that they have a number of possible columns for mapping.

## Part 2 - Mapping
In order to map the Black population in Missouri correctly, we need to decide on how we want to normalize our map. We have three column choices in this lab:

  1. `black_den` is the number of African American residents per square kilometer
  2. `black_pct` is the percentage of African American residents in each county
  3. `blackr_` is the per capita rate per 1,000, so, the number of Black residents per 1,000 total residents
  
Any of these three are appropriate for this map!

```{r, map-1}
p1 <- ggplot() +
  geom_sf(data = mo, mapping = aes(fill = black_pct)) +
  scale_fill_viridis(name = "% African American") +
  labs(
    title = "Black Population by Missouri County",
    subtitle = "2010 U.S. Census",
    caption = "Data via the U.S. Census Bureau \n Map by Christopher Prener, PhD"
  )

p1
```

If we map using percent, we see that St. Louis City has the highest percent of African American residents. Other counties with relatively high percentages are St. Louis County, Jackson County (the main part of Kansas City), and the Bootheel counties in southeastern MO.

If we map using density (not shown), the only counties that stand out are St. Louis City, which has the highest density of African American residents, as well as St. Louis County and Jackson County (the main part of Kansas City).

If we map using per rate per 1,000, we see that St. Louis City has the highest rate of African American residents. Other counties with relatively high rates are St. Louis County, Jackson County (the main part of Kansas City), and the Bootheel counties in southeastern MO.

Finally, we'll save our map to the `results/` folder. Remember that you'll have had to create this ahead of time!

```{r, save-map-1}
ggsave(plot = p1, filename = here("results", "mo_african_americans.png"))
```

Now our map has been saved!
